Stateless Model Checking with Data-Race Preemption Points

### 2015-11-12 pldi-16/evaluation.tex
some rephrasing

@@

% TODO: You should introduce this in the narrative as a key contribution way earlier.
\subsection{Finding nondeterministic data-race candidates}
\label{sec:eval-falseneg}
Some memory accesses may be hidden in a control flow path that requires a nondeterministic preemption to be executed.
In such cases, a single-pass dynamic data-race detector
%could not achieve the coverage necessary
would fail
to identify a racing access pair as a candidate to begin with.
%
We counted how many such data-races, used as PPs, led to \quicksand~finding new bugs,
thereby making them {\em false negatives} of the single-pass approach.
% TODO: Put a figure here giving an example of where e.g. a data race only shows up during the contention path of a mutex.
We classified each data-race candidate according to whether \landslide~had reported them during the first interleaving, before any backtracking or preempting: if so, they were {\em deterministic data races} (hence could be found by single-pass).

@@

				% TODO: Argue:
				% It is fair to compare multiple pass DR-analysis under Landslide against just a single execution because prior work DR detectors, being not integrated with a MC, are not intended to uncover different results under subsequent runs.
				% Define a "stress tester" as a class of bug detectors where they are intended to [...]
				% Maybe say: Do there exist any stress-tester DR detectors where they are intended to produce new results under reruns?

@@

% TODO: restructure this paragraph to account for RaceFuzzer/Portend. Be like, "Yes they explore multiple schedules, but only AFTER finding a race." And/or make this argument in prior work
One might also wonder: Why is it fair to compare the data race bugs \quicksand~finds (10 CPU-hours)
against the candidates found by a prior work's single execution (less than a minute)?
We argue this comparison is meaningful because prior work data-race tools, being not integrated with a MC,
are not intended to discover new data race candidates under subsequent runs.
Running a data-race tool repeatedly for 10 CPU-hours could potentially uncover some of these false negatives at random,
but stress testing's comparative problem with achieving reliable coverage is already well-understood
% TODO: 1 more good citation here
\cite{chess-icb,gambit}.
% TODO: read gambit paper
%One could run a data-race tool repeatedly for 10 CPU-hours, but the advantage of stateless MC over stress testing is already well-understood.
% TODO TODO TODO get a big citation for teh above sentence.

@@

% TODO Rename to "Limitations and Future Work"
\section{Discussion}
\label{sec:future}

@@

In some cases, this repeated work can cause Iterative Deepening to be slower than SSS-MC to find certain bugs,
for example, if both {\tt mutex\_lock} and {\tt mutex\_lock} preemptions together would expose a bug, but not either alone.
Similarly, when pursuing total verification (\sect{\ref{sec:totalverif}}),
if the state space resulting from preempting on every instruction could be completed,
SSS-MC will achieve that verification faster, as Iterative Deepening will test many subsets of data-race PPs first.
Predicting whether each still-untested interleaving will uncover new PPs is not straightforward,
but we believe future implementations of Iterative Deepening could incorporate cross-job memoization
to prune some or all of this repeated work.

>>>>

% TODO: Make a CDF graph of how many tests got total verification here.

% TODO: Figure out how to visualize how many unique data race PPs were verified, and how many remained as unclassified false positives

@@

% TODO: You should introduce this in the narrative as a key contribution way earlier.
\subsection{Finding nondeterministic data-race candidates}
\label{sec:eval-falseneg}
Some memory accesses may be hidden in a control flow path that requires a nondeterministic preemption to be executed.
In such cases, a single-pass dynamic data-race detector
%could not achieve the coverage necessary
would fail
to identify a racing access pair as a candidate to begin with.
%
We counted how many such data-races, used as PPs, led to \quicksand~finding new bugs,
thereby making them {\em false negatives} of the single-pass approach.
We classified each data-race candidate according to whether \landslide~had reported them during the first interleaving, before any backtracking or preempting: if so, they were {\em deterministic data races} (hence could be found by single-pass).

@@

@@

At first, it might seem unfair to compare the data race bugs \quicksand~finds (10 CPU-hours)
against the candidates found by a prior work's single execution (less than a minute).
We argue this comparison is meaningful because prior work data-race tools %, being not integrated with a MC,
are not intended to discover new data race candidates under subsequent runs. %before discovering any initial candidates.
Running a data-race tool repeatedly for 10 CPU-hours could potentially uncover some of these false negatives at random,
but stress testing's comparative problem with achieving reliable coverage is already well-understood
% TODO: 1 more good citation here
\cite{chess-icb,gambit}.
%Prior work data-race tools \cite{portend} could be configured to have the same sort of feedback loop as Iterative Deepening,
%i.e., discovering transitively-reachable data races while testing initial ones,
%although that would still not simultaneously be able to provide total verifications.
% TODO: read gambit paper

@@

\section{Discussion}
\label{sec:future}

@@

In some cases, this repeated work can cause Iterative Deepening to be slower than SSS-MC to find certain bugs,
for example, if both {\tt mutex\_lock} and {\tt mutex\_lock} preemptions together would expose a bug, but not either alone.
Similarly, when pursuing total verification (\sect{\ref{sec:totalverif}}),
if the state space resulting from preempting on every instruction could be completed,
an SSS-MC tool such as \cite{spin} might achieve that verification faster,
as Iterative Deepening will test many subsets of data-race PPs first.
Predicting whether each still-untested interleaving will uncover new PPs is not straightforward,
but we believe future implementations of Iterative Deepening could incorporate cross-job memoization
to prune some or all of this repeated work.

### 2015-11-12 pldi-16/intro.tex
some rephrasing

This paper's contributions are as follows:
\begin{enumerate}
	\item Iterative Deepening, a new technique for combining data-race analysis with stateless model checking, and \quicksand, an open source implementation of the technique;
	\item A proof of convergence, which shows that fully testing every discovered data-race preemption point is equivalent to testing all possible thread schedules, should that be possible in the given CPU budget;
	\item A new tactic for eliminating one class of false-positive data races,
		which cannot soundly be used in a single-pass analysis,
		but which we prove correct when used with Iterative Deepening;
		%, unsound in single-pass analysis but which we prove sound when used with Iterative Deepening;
	%\item Techniques for detecting and flattening cyclic state spaces resulting from ad-hoc while-loop synchronization % TODO: is there actually room for this in the paper?
	\item A large evaluation in which \quicksand~compares favorably to stand-alone data-race detection and stateless model checking approaches.
\end{enumerate}

>>>>

This paper's contributions are as follows:
\begin{enumerate}
	\item Iterative Deepening, a new technique for combining data-race analysis with stateless model checking, and \quicksand, an open source implementation of the technique;
	\item A proof of convergence, which shows that should it be possible in the given CPU budget,
		fully testing every discovered data-race preemption point is equivalent to testing all possible thread schedules;
	\item A new tactic for eliminating one class of false-positive data races,
		which cannot soundly be used in a single-pass analysis,
		but which we prove correct when used with Iterative Deepening;
		%, unsound in single-pass analysis but which we prove sound when used with Iterative Deepening;
	%\item Techniques for detecting and flattening cyclic state spaces resulting from ad-hoc while-loop synchronization % TODO: is there actually room for this in the paper?
	\item A large evaluation in which \quicksand~compares favorably to stand-alone data-race detection and stateless model checking approaches.
\end{enumerate}
